### 2022-08-04 redaction/EHMOS_ESM1.Rmd
initial commit

---
title: ""
classoption: a4paper
output:
  pdf_document:
  word_document: default
  documentclass: article
fontsize: 12pt
header-includes:
- \usepackage{float}
- \usepackage{lscape}
- \usepackage{amsmath}
- \usepackage{lineno}
#- \linenumbers
#- \usepackage{setspace}\doublespacing
csl: journal-of-biogeography.csl
bibliography: Paper_EHMOS_cut.bib
---

```{r setup, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
# set chunk default options for the document
knitr::opts_chunk$set(echo=FALSE, 
                      message=FALSE, 
                      warning=FALSE, 
                      fig.pos = 'H',
                      fig.align = 'center')
```

## Appendix S1: Description of EHMOS' parameters

In the occupancy model of EHMOS (see Eq. (3) in the main text), the four parameters ($\alpha$, $\theta$, $\delta$ and $\tau$) represent ecological descriptors of the species-environment relationship (Fig. S1):

* $\alpha$ corresponds to the maximum probability of occurrence, $\psi_{max}$, on the logit scale. Indeed, the maximum probability of occurrence is reached when gradient value is equal to the optimum, i.e. we have $\psi_j = \psi_{max}$ when $X_j = \theta$. Thus, Eq. (3) in the main text becomes:

\begin{equation}
  log(\frac{\psi_{max}}{1-\psi_{max}}) = \alpha \iff \psi_{max}  =   \frac{1}{1+ \exp(-\alpha)};
\end{equation}

* $\theta$ is the species optimum, the environmental value at which species reached its maximum probability of occupancy;

* $\delta$ is the species shift between two optima;

* $\tau$ represents the environmental tolerance of a species, i.e. the gradient range that a species can occupy. It can be related to the width, $\omega$, of the species response curve at a specified occupancy probability threshold, $p_{\omega}$, by:

\begin{equation}
  \omega = 2 \times \tau \sqrt{2 \times (\alpha - logit(p_{\omega}))} \iff \tau = 0.5 \times  \frac{\omega}{\sqrt{2 \times (\alpha - logit(p_{\omega}))}}
\end{equation}
with a $p_{\omega}$ usually specified at 0.05 to estimate the suitable range of a species [@michaelis2017].

Those four EHMOS parameters can also be derived from cGLMM coefficient estimates [@jamil2014]:

\begin{align*}
  \alpha &= \beta_0 - \frac{{\beta_1}^2}{4 \beta_2} \\
  \theta &= - \frac{\beta_1}{2 \beta_2} \\
  \delta &= - \frac{\beta_1}{2 \beta_2} - (-\frac{\beta_1 + \beta_4}{2 (\beta_2 + \beta_5)}) \\
  \tau &= \sqrt{-\frac{1}{2 \beta_2}}
\end{align*}

```{r schema_param, fig.cap = "", fig.align='center'}

img <- png::readPNG(here::here("results", "figs", "schema_parameters.png"))
grid::grid.raster(img)

```

Figure S1.1: Schematic representation of ecological parameters that describes species-environment relationship.

## References
<div id="refs"></div>