### 2022-08-04 redaction/EHMOS_ESM2.Rmd

---
title: ""
classoption: a4paper
output:
  pdf_document:
  word_document: default
  documentclass: article
fontsize: 12pt
header-includes:
- \usepackage{float}
- \usepackage{lscape}
- \usepackage{amsmath}
- \usepackage{lineno}
#- \linenumbers
#- \usepackage{setspace}\doublespacing
csl: journal-of-biogeography.csl
bibliography: Paper_EHMOS_cut.bib
---

## Appendix S2: Description of performance metrics used in the simulation study

The four performance metrics used (bias, RMSE, interval coverage and interval score) are computed for each model at different levels (species, scenario, overall). First, for each species $i$ included in a particular scenario $s$, we computed average perfomance metrics over the thirty replications:

* Bias of species $i$ in scenario $s$ based on method $m$ is defined as the average of differences between absolute^[Absolute values are took because true shifts could either be positive or negative, thus a negative bias always represents an underestimation of the true shif magnitude whatever its direction.] estimates based on method $m$ and absolute true shift over the thirty replications:

$$Bias_{i,s,m} = \frac{1}{R} \sum_{r=1}^{R}|\hat{\delta}_{i,s,m,r}| - |\delta_{i,s}| \text{ ;}$$

* The RMSE at species-specific level is defined by:

$$RMSE_{i,s,m} = \sqrt{\frac{1}{R}\sum_{r=1}^{R} (\hat{\delta}_{i,s,m} - \delta_{i,s})^2} \text{ ;}$$

* The interval coverage for a species in a particular scenario is the ratio between the number of replciations in which confidence interval estimate contain the true shift and the total number of replications;

* The interval score of species $i$ in scenario $s$ based on method $m$ is defined as the average of interval scores obtained at each replication $r$. At the replication level, interval score is computed as follows:

\begin{equation*}
IS_{i,s,m,r}(l_{i,s,m,r}, u_{i,s,m,r}; \delta_{i,s}) = 
\begin{cases}
(u_{i,s,m,r}-l_{i,s,m,r}) + \frac{2}{\alpha}(l_{i,s,m,r} - \delta_{i,s}) &\text{if $\delta_{i,s} < l_{i,s,m,r}$}\\
(u_{i,s,m,r}-l_{i,s,m,r}) &\text{if $u_{i,s,m,r} > \delta_{i,s} > l_{i,s,m,r}$}\\
(u_{i,s,m,r}-l_{i,s,m,r}) + \frac{2}{\alpha}(\delta_{i,s} - u_{i,s,m,r}) &\text{if $\delta_{i,s} > u_{i,s,m,r}$}
\end{cases}
\end{equation*}

where $\alpha$ is the confidence level of the interval, here set at 0.05, $l_{i,s,m,r}$ and $u_{i,s,m,r}$ are the lower and upper limits of the interval estimated by model $m$ for species $i$ in scenario $s$ and replicate $r$.

At the scenario level, metrics are averaged over the twenty species:
$$\overline{Bias}_{s,m} = \frac{1}{N} \sum_{i=1}^{N} Bias_{i,s,m}$$
$$\overline{RMSE}_{s,m} = \frac{1}{N} \sum_{i=1}^{N} RMSE_{i,s,m}$$
$$\overline{Coverage}_{s,m} = \frac{1}{N} \sum_{i=1}^{N} Coverage_{i,s,m}$$
$$\overline{IS}_{s,m} = \frac{1}{N} \frac{1}{R} \sum_{i=1}^{N} \sum_{r=1}^{R} IS(\hat{l}_{i,s,m,r}, \hat{u}_{i,s,m,r}; \delta_{i,s})$$