### 2022-11-04 README.Rmd
Commit before publication

---
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below.
Title <- "A new method to explicitly estimate the shift of optimum along gradients in multispecies studies"
```

This repository contains the data and code for our paper submitted to _Journal of Biogeography_:

> B. Mourguiart, B. Liquet, K. Mengersen, T. Couturier, J. Mansons, Y. Braud, A. Besnard , (under revision). _`r Title`_.